---
title: "EDA"
output: 
  html_document:
    df_print: paged
  pdf_document: default
---

Reading data

```{r}
library('dplyr',quietly = TRUE)
library('ggplot2',quietly = TRUE)
library('gridExtra',quietly = TRUE)
library('grid',quietly = TRUE)
library("rgdal",quietly = TRUE)
library("choroplethrMaps",quietly = TRUE)
library("choroplethr",quietly = TRUE)
library('ClustOfVar', quietly = TRUE)
library('corrplot', quietly = TRUE)
library('cluster', quietly = TRUE)

#data = read.csv("C:/Users/mua377n/Downloads/archive/Loan_status_2007-2020Q3.gzip")
data = data %>% select (-id)
print(dim(data))
colnames(data)
```

missing values plot for different variables

```{r}
missing_values = apply(is.na(data),2, sum )/nrow(data)
barplot(missing_values,legend.text = 'variables', main='NAs for different variables')
```

treating loan_status

```{r}
loan_status_keep = c(unique(data$loan_status)[1],
                     unique(data$loan_status)[2],
                     unique(data$loan_status)[6],
                     unique(data$loan_status)[7],
                     unique(data$loan_status)[9],
                     unique(data$loan_status)[10])
data_new = data[data$loan_status %in% loan_status_keep,]
remove(data)
ggplot(data = data_new)+geom_bar(mapping = aes(x = loan_status))
```

dropping temporal variables

```{r}
drop_colums = c('last_credit_pull_d','last_fico_range_high','last_fico_range_low','total_pymnt',
                        'total_pymnt_inv','recoveries','collection_recovery_fee','out_prncp','out_prncp_inv',
                        'total_rec_prncp','total_rec_int','last_pymnt_d','last_pymnt_amnt','next_pymnt_d',
                        'total_rec_late_fee','hardship_flag','hardship_amount',
                        'orig_projected_additional_accrued_interest','hardship_payoff_balance_amount',
                        'hardship_last_payment_amount','debt_settlement_flag','hardship_type','hardship_reason',
                        'hardship_status','deferral_term','hardship_start_date','hardship_end_date','payment_plan_start_date',
                        'hardship_length','hardship_dpd','hardship_loan_status','installment','pymnt_plan',                        'acc_now_delinq','url','earliest_cr_line','last_fico_range_low','last_fico_range_high','policy_code','zip_code',
                      'sec_app_fico_range_low','sec_app_fico_range_high','sec_app_earliest_cr_line','fico_range_low', 'fico_range_high',
                     'annual_inc_joint','dti_joint','verification_status_joint', 'revol_bal_joint','sec_app_inq_last_6mths',
                     'sec_app_mort_acc','sec_app_open_acc','sec_app_revol_util','sec_app_open_act_il',
                      'sec_app_num_rev_accts','sec_app_chargeoff_within_12_mths','sec_app_collections_12_mths_ex_med')

data_new = data_new[,!names(data_new)%in% drop_colums]
# data_new = data_new %>% select (-one_of(drop_colums)
#stripping % from charater type and converting to numeric
data_new$int_rate = as.numeric(sub("%", "", data_new$int_rate))
data_new$revol_util = as.numeric(sub("%", "", data_new$revol_util))
data_new$id = as.integer(data_new$id)
```

visualizing grade

```{r}
ggplot(data = data_new)+geom_bar(mapping = aes(x = grade))
```

visualizing home ownership

```{r}
ggplot(data = data_new)+geom_bar(mapping = aes(x = home_ownership))
```

visualizing term

```{r}
ggplot(data = data_new)+geom_bar(mapping = aes(x = term))
```

visualizing emp_length

```{r}
ggplot(data = data_new)+geom_bar(mapping = aes(x = emp_length))
```

creating target Y variable coFlag

```{r}
data_new$coFlag = ifelse (data_new$loan_status %in% c(loan_status_keep[1], loan_status_keep[5], loan_status_keep[3]), 0, 1)
```

visualizing Y variable

```{r}
ggplot(data = data_new)+geom_bar(mapping = aes(x = coFlag))
```

```{r}
ggplot(data_new, mapping= aes(coFlag))+geom_bar(mapping = aes(color= loan_status))
```

```{r}
ggplot(data_new, mapping= aes(coFlag))+geom_bar(mapping = aes(color = grade))
```

```{r}
ggplot(data_new, mapping= aes(coFlag))+geom_bar(mapping = aes(color = emp_length))
```

```{r}
ggplot(data_new, mapping= aes(coFlag))+geom_bar(mapping = aes(color = home_ownership))
```

```{r}
ggplot(data_new, mapping= aes(coFlag))+geom_bar(mapping = aes(color = purpose))
```

```{r}
ggplot(data_new, mapping= aes(coFlag))+geom_bar(mapping = aes(color = verification_status))
```

seperating continuous and categorical variables

```{r}
all_vars <- unlist(lapply(data_new,class))

#continuous variable
cont_var = all_vars[all_vars=='integer' | all_vars == 'numeric']
cont_var = cont_var[c(1:71)]

#categorical variable
cat_var = all_vars[all_vars=='character']
```

```{r}
data_cont_var = data_new[,names(data_new) %in% names(cont_var)]
summary(data_cont_var)
```

```{r}
loan_by_state <- data_new %>% 
                 group_by(addr_state) %>%
                 summarize(`Total Loans ($)` = sum(loan_amnt)/1e6) %>%
                 arrange(desc(`Total Loans ($)`))
colnames(loan_by_state) <- c("region","value")
top4_states <- round(100*sum(loan_by_state$value[1:4])/sum(loan_by_state$value),1)
data("state.regions")
loan_by_state$region <- sapply(loan_by_state$region, function(state_code) {
  inx <- grep(pattern = state_code, x = state.regions$abb)
  state.regions$region[inx]
})
print(state_choropleth(loan_by_state, title = "           Total Loan Volume by State - Millions $"))
```

```{r}
ggplot(data = data_new) + geom_count(mapping = aes(x = grade, y = emp_length))
```

```{r}
ggplot(data = data_new) + geom_count(mapping = aes(x = grade, y = home_ownership))
```

```{r}
ggplot(data = data_new) + geom_count(mapping = aes(x = grade, y = purpose))
```

```{r}
ggplot(data = data_new) + geom_count(mapping = aes(x = grade, y = term))
```

```{r}
ggplot(data=data_new, mapping = aes(int_rate)) + geom_freqpoly(mapping = aes(color = loan_status), binwidth =1)
```

correlation heatmap

```{r}
m = data_new %>% mutate_if(is.character,as.factor)
m = m%>% mutate_if (is.factor, as.numeric)
m = m[,-1]  
m = m[,-86]
m = cor(m)
print(m[c(1:17),c(1:17)])
```
```{r}
corrplot(m[c(1:17),c(1:17)], method="color", tl.col= 'black', is.corr = FALSE)
```
variable clustering / produces a dendogram

```{r}
missing_values_new = sort(apply(is.na(data_new),2, sum )/nrow(data_new),TRUE)
barplot(missing_values_new,legend.text = 'variables', main='Removing variables with more than 10% values missing')
drop_col_missing = names(missing_values_new[c(1:20)])
data_new = data_new[,!names(data_new) %in% drop_col_missing]

#xcat = data_new[,names(data_new) %in% names(cat_var[-1])]
#xcont = data_new[,names(data_new) %in% names(cont_var)]
#tree =kmeansvar(xcont, xcat, init = 4)
#plot(tree)
```
